Type-Safe Modular Parsing

### 2017-01-12 paper/sections/ObjectAlgebras.tex
Section 4

\huang{to polish}As illustrated above, one intuition of using Object Algebras is to separate data structures from behaviors, so as to modularize the design, and both parts are even more modular and reusable. We can simply add new variants corresponding to the new cases in the grammar, and realize different operations on structures, with both code reuse and separate compilation supported.

Another advantage of using Object Algebras is that we benefit from using generics. If the syntax contains multiple sorts, we can distinguish them by different type parameters. Thus fewer classes are defined and code is simplified. As an example, we extend the language of literals, additions and variables by introducing types as a new sort. Now it has two sorts in the syntax, types and expressions. We certainly do not want to mix them together.\\

@@

When extending the Object Algebra interface, pretty-printing operation and parser, we use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types. They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.

\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. For the same reason,
we use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, recursively. \inlinecode{pT} and \inlinecode{pE} are actually the parsers for recursion.

From the code we observe that the multi-sorts of Object Algebra interface
exactly express multiple syntax. The isolation for different syntax
is guaranteed by the type system. Now it is straightforward to show some client code
for the new parser:

\lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
It works just as expected to parse and pretty-print the result.

>>>>

\begin{comment}
As illustrated above, using Object Algebras separates data structures from behaviors, thus enabling more modularity and reuse. New language constructs correspond to the new cases in the algebra. Different operations
 on structures, with both code reuse and separate compilation supported.
\end{comment}

Another advantage of using Object Algebras is that is is easy to 
model multi-sorted languages. If the syntax contains multiple sorts, we can distinguish them by different type parameters. As an example, we extend the language of literals, additions and variables by introducing types as a new kind of type argument, and we add lambda abstractions to expressions. Now the language has two sorts in the syntax, types and expressions:\\

@@

\begin{figure}
\lstinputlisting[linerange=31-59]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX
\caption{Code for parsing a language with multi-sorted syntax for types and expressions.}
\label{fig:multi}
\end{figure}

Figure~\ref{fig:multi} illustrates the corresponding Scala code 
that extends the Object Algebra interface, pretty-printing operation and parser. We use two type parameters \inlinecode{E} and \inlinecode{T} for expressions and types. They guarantee that invalid terms such as \inlinecode{int + int} will be rejected by the parser.
Besides lexing, the trait \inlinecode{TypedLamOAParser} also introduces parsers for types, and the new case for expressions. 
We use \inlinecode{pTypedLamT} and \inlinecode{pTypedLamE} as copies of current \inlinecode{pT} and \inlinecode{pE}, due to the issue 
with \lstinline{super} in Scala (see discussion in Section~\ref{subsec:parsingwithoa}). \inlinecode{pT} and \inlinecode{pE} are the parsers used for performing recursion.

From the code we observe that the multi-sorts of Object Algebra interface
express multiple syntax. The isolation for different syntax
is guaranteed by the type system. The following illustrates some client code
for the new parser:

\lstinputlisting[linerange=63-65]{../Scala/Parser/src/PaperCode/Sec4OA/Code4.scala}% APPLY:linerange=OVERVIEW_OA_MULTI_SYNTAX_CLIENT
The code works as expected to parse and pretty-print the result.

Note that, similarly to the approach in Section~\ref{sec:inheritance} independent extensibility is also supported via multiple trait inheritance. 
Since independent extensibility is achieved using essentially the same technique as in Section~\ref{sec:inheritance} we omit that code here.
